---
title: "End-to-end research pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{End-to-end research pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette is the monorepo showcase: one research question, answered
by four packages working together. Each layer depends on the one before
it, and the build system resolves that chain automatically.

| Layer | Package | Role | Internal deps |
|-------|---------|------|---------------|
| 1 | `uofgcore` | validation helpers, compiled C interop | — |
| 2 | `uofgstats` | rolling statistics, models, data contracts | `uofgcore` |
| 3 | `uofgdata` | reproducible synthetic market data | `uofgstats` |
| 4 | `uofgbacktest` | causal backtesting engine | `uofgcore`, `uofgdata`, `uofgstats` |

```{r setup, message = FALSE}
library(uofgbacktest)
```

## The question

Does a fast-over-slow moving-average crossover earn its transaction
costs on daily data? We answer it on synthetic data so the result is
reproducible everywhere, forever.

## 1. Generate data (uofgdata)

```{r}
mkt <- uofgdata::simulate_ohlcv(n_days = 500, seed = 42, vol = 0.01)
head(mkt, 3)
```

The series satisfies the data contract by construction:

```{r}
uofgstats::validate_ohlcv(as.data.frame(mkt))
```

## 2. Time-based split (backtesting hygiene)

```{r}
prices <- as.numeric(mkt$Close)
splits <- time_split(n = length(prices), train_frac = 0.7)
stopifnot(max(splits$train) < min(splits$test))
```

The signal parameters would be tuned on `train`; `test` is touched once,
at the end.

## 3. Backtest (uofgbacktest)

The engine enforces causality in code: a signal that peeks ahead is
refused, not merely warned about.

```{r}
bt_train <- run_backtest(
  prices[splits$train],
  ma_signal,
  cost_per_trade = 0.001
)
bt_train
```

## 4. Evaluate out of sample

```{r}
bt_test <- run_backtest(
  prices[splits$test],
  ma_signal,
  cost_per_trade = 0.001
)
summary(bt_test)
```

## 5. Sanity-check the numerics

`rolling_mean()` powers the signal. Its two methods agree on this
well-scaled series — and the test suite verifies both against
arbitrary-precision (`Rmpfr`) references, so this agreement is enforced,
not hoped for:

```{r}
stopifnot(all.equal(
  uofgstats::rolling_mean(prices, 20, method = "exact"),
  uofgstats::rolling_mean(prices, 20, method = "fast")
))
```

## Takeaway

Every layer did one job: `uofgcore` validated, `uofgstats` computed and
contract-checked, `uofgdata` made it reproducible, `uofgbacktest`
enforced methodology. Because they live in one monorepo, all four were
built, tested, checked, and versioned together by a single
`Rscript scripts/build-all.R`.
