---
title: "Backtesting hygiene"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Backtesting hygiene}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A backtest is a numerical experiment, and experiments are only as
trustworthy as their discipline. This vignette collects the hygiene
rules this monorepo's tooling is designed to enforce, with worked
examples.

```{r setup, message = FALSE}
library(uofgstats)
library(xts)
withr::local_seed(2026)
```

## 1. Validate the data before trusting any result

Corrupted bars (High < Low, missing volumes, mis-sorted timestamps)
produce confident nonsense. Reject them before analysis:

```{r}
prices <- data.frame(
  Open   = c(100, 101, 102),
  High   = c(102, 103, 104),
  Low    = c(99, 100.5, 101),
  Close  = c(101, 102, 103.5),
  Volume = c(1000, 1200, 900)
)
validate_ohlcv(prices)
```

## 2. Causality: never let the future leak

Every rolling function here is **trailing**: window `i` uses only data
up to and including `i`. Truncating the future must never change past
values — this is the look-ahead test, and you should apply it to every
signal you write:

```{r}
x <- c(1, 2, 3, 4, 5, 6, 7)
stopifnot(identical(
  rolling_mean(x, 3)[1:3],
  rolling_mean(x[1:5], 3)
))
```

If a custom signal passes this test, it cannot peek ahead.

## 3. Numerical accuracy at scale

Rolling statistics over long histories accumulate floating-point drift
if implemented with running sums. Compare the two methods on an
adversarial series, and check against arbitrary-precision arithmetic
in your own analyses with `Rmpfr`:

```{r}
x <- c(rep(1e16, 2), rep(1, 8))
rolling_mean(x, 3, method = "exact") # exact by construction
```

Use `method = "fast"` for speed on well-scaled data; the default
protects you when magnitudes are extreme.

## 4. Time-based splits only

Random train/test splits leak the future into the past. Split by time,
and assert that the split is clean:

```{r}
dates <- seq(as.Date("2025-01-01"), by = "day", length.out = 100)
prices <- xts(x = cumprod(1 + rnorm(100, 0.0004, 0.01)), order.by = dates)
colnames(prices) <- "strategy"

split_date <- as.Date("2025-03-01")
cut <- which(zoo::index(prices) > split_date)[1L]
train <- seq_len(cut - 1L)
test <- seq.int(cut, length(prices))
stopifnot(!any(train %in% test), max(train) < min(test))
```

## 5. Costs compound: model them

A strategy with mean daily gross return $\mu_g$ and proportional cost
$c$ per position change has net mean approximately
$\mu_g - c \cdot \mathbb{E}|\Delta \text{position}|$ per rebalance.
Even tiny costs dominate small edges:

```{r}
gross <- to_returns(as.numeric(prices))
turnover <- abs(diff(signals <- rep(c(1, 0, 1), length.out = length(gross))))
cost_per_trade <- 0.001 # 10 bps
net <- gross * signals - cost_per_trade * turnover
c(gross_mean = mean(gross), net_mean = mean(net))
```

## 6. Reproducible randomness

- Fix seeds locally with `withr::with_seed()` so helpers never change
  global state.
- For parallel Monte Carlo, use L'Ecuyer streams so results are
  reproducible regardless of worker count:

```{r}
withr::with_seed(42, x <- rnorm(3))
withr::with_seed(42, y <- rnorm(3))
stopifnot(identical(x, y))
```

```{r, eval = FALSE}
# parallel-safe seeding pattern:
RNGkind("L'Ecuyer-CMRG")
set.seed(2026)
streams <- parallel::nextRNGStream(123)
```

## Summary checklist

1. `validate_ohlcv()` before analysis.
2. Every signal passes the truncation (look-ahead) test.
3. `method = "exact"` when magnitudes are extreme; verify with `Rmpfr`.
4. Time-based splits, never random ones.
5. Costs modeled per rebalance.
6. Seeds via `withr`; parallel via L'Ecuyer streams.
