---
title: "Getting started with uofgstats"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with uofgstats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette walks through the two main entry points of uofgstats:
`rolling_mean()` for moving-window summaries and `fit_linreg()` for
linear models with a tidy coefficient table. It also demonstrates the
package conventions this monorepo follows.

```{r setup}
library(uofgstats)
```

## Rolling means

`rolling_mean()` computes trailing means in O(n) using a cumulative-sum
algorithm. Input validation is shared with `uofgcore`, so bad input
fails with a consistent error style across the whole monorepo:

```{r}
rolling_mean(c(1, 2, 3, 4, 5), 2)
```

```{r, error = TRUE}
rolling_mean(c(1, 2), 5)
```

## Linear models

`fit_linreg()` wraps `stats::lm()` in a light S3 class. Printing it
gives a compact summary, and `tidy_linreg()` returns coefficients as a
plain data frame (no tidyverse dependency required):

```{r}
fit <- fit_linreg(y ~ x, data.frame(x = 1:5, y = c(2.1, 3.9, 6.2, 8.1, 9.8)))
tidy_linreg(fit)
```

Additional arguments are forwarded to `stats::lm()`, including
non-standard-evaluation arguments like `weights`:

```{r}
weighted <- data.frame(x = 1:3, y = c(1, 2, 9), w = c(1, 1, 10))
tidy_linreg(fit_linreg(y ~ x, weighted, weights = w))
```

## Validation

Helpers from `uofgcore` are re-exported nowhere; call them through the
namespace. Validation errors always name the offending argument:

```{r, error = TRUE}
uofgcore::validate_numeric("not numeric")
```
